Cross-sectoral conservation of the Indian Ocean

### 2022-03-09 Msc_Thesis_Lea_v2.Rmd
Update Msc_Thesis_Lea_v2.Rmd
added prioritizr problem solving and plotting

---
title: "Cross-sectoral conservation of the Indian Ocean"
author: "Lea Fourchault"
date: "22/02/2022"
output: html_document
---
<style type="text/css">

@@

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

@@

# Table of Contents
# Introduction
# Materials & Methods
## Workflow with Prioritizr 
## Overview of data  & layers
## Planning region 
## Conservation features & targets
### IBAs
### IMMAs
### EBSAs
### Deep-sea features : seamounts, ridges, plateaus, vents 
### Targets dataframe
## Cost layer
### CFC
### PMN
### Exploration areas
### Reserved areas
### Shipping
### Fishing *use cost layer already created?*
# Prioritizr problems
### Basic run (only Cons features)
### Cons features, fishing
### Cons features, mining (lock-out expl  & res areas; assign cost for PMN areas)
### Cons features, fishing + mining 
### Cons features, shipping 
# Discussion
## Potential improvements
# Conclusion

+ how to group all files (ex: all EBSAs + IBAs + IMMAs + vents + seamounts...) into one, to create ConsFeatures? Or is it enough to have columns next to PUs? RESOLVED: yes, stacking = creating columns next to PUs.
+ should we use Sharma (2011) + CRU Consulting (2019) to assign USD values to PMN areas? see PMN subsection. RESOLVED, yes, can use for this project (need to cross-check if want to publish)
+ shipping raster : cells have certain values (# of ships recorded?? I can't figure from the attached files, metadata is confusing), could allow to infer USD value of route (# of ships * approx cost of fuel for average cargo to go through 1 PU, which is 1000sqkm)? So we wouldn't try to compute the cost of deviating from any given route, but rather compute the price for each ship to go through 1 PU * the number of ships going through that PU. It sounds fairly logical to me, but maybe it's not? RESOLVED: check 2008 metada, run prio problem with only shipping
+ fishing cost layer : there is a ready-made fx in SpatPlan repo (thanks Jase & Tin), should we use it for this project? see fishing subsection (but issues with plotting  as you can see). RESOLVED: yes, use fx from Jase & Tin. Change scale to remove outliers/crop to planning region.
+ I'm using echo = FALSE, message = FALSE, warning = FALSE, results = "hide" and still can't get rid of some coding output in the .html. RESOLVED: cut up chuncks of code with separate ones for plots
+ I st_to_sf the vents from .csv with PUs' crs but have some issues intersecting the points with polygs, will look more into it. RESOLVED: use st_contains after crs = 4326 & st_transform(cCRS)
+ add list of species as cons_feats: EN, VU etc (Ant's email to mining colleague)? RESOLVED: fine-tune later
+ shipping raster (Halpern et al., 2015) is too heavy to be converted to polygon, what could be other methods (not using as.polygon)? RESAMPLE if resolution <1°. Resolution = 934m*934m (! 1° at equator = 111km = 60 nm; [convert](https://www.johndcook.com/how_big_is_a_degree.html); resolution too fine so need resampling). RESOLVED: no resampling, overlay raster with PUs (exactextractr) to attribute mean raster cells value to PUs (no weighted mean because raster cells are all equal-area because Mollweide proj)
+ how to remove EEZs from PUs? RESOLVED: find EEZ shp and spatial predicates, inverse overlay with Bndry, recreate PUs
+ setting  targets. RESOLVED: 30% for rpz area, 70% for important/rare feature; create dataframe with 1 column: features, 1 column: targets (match order)
+ cross inactive vents & PUs - done

Current questions & issues:

+ ongoing - pricing deep-sea mineral resources (I am looking for more papers especially regarding CFC)
+ check research questions
+ overlay shipping & PUs + plot with log10?
+ how to assign price to mining shps / create cost layer?
+ assigning targets - Cons_feats tibble created

@@

+ PUs = ABNJ only
+ cons feats = IMMAs, IBAs, chosen EBSAs, vents (active + inactive), seamounts, plateaus; no aquamaps
+ mining cost = value/t * density * area, discount rate and other improvements to be determined later (if any), USD
+ fishing cost = removed fish mass * fish value (layer created by Jase), USD
+ shipping cost = mean of shipping raster cells' values for each PU, not USD (intensity)
+ targets = 30% for rpz area, 70% for important

@@

@@

**Possible research questions**

Ranked from most general to most precise, can be same question with different wording (some are in passive form, I know it's better to use active but some sound weird when saying 'we'?)

1. how can marine spatial planning adapt to a time of increased anthropogenic pressures?
2. how can the interests of multiple industrial stakeholders be incorporated into MPA design?
3. how can value be estimated in the context of data-poor industries?
4. which areas are most valuable to industrial stakeholders in the Indian Ocean?
5. how can we design MPAs that minimize trade-offs with industrial stakeholders in the Indian Ocean?
6. how can we design MPAs that maximize biodiversity conservation across depth layers in the Indian Ocean?
7. how can we quantify/estimate trade-offs amongst industries in the Indian Ocean?
8. how can we quantify/estimate trade-offs between industries and conservation in the Indian Ocean? and,
9. how can such trade-offs with industrial stakeholders be minimized when designing MPAs?

@@

blue habitats **do not include hydro vents**, so will use Beaulieu & Szanfranski (2020) [link](https://doi.pangaea.de/10.1594/PANGAEA.917894?format=html#download)

@@

cCRS <- "+proj=robin +lon_0=0 +x_0=0 +y_0=0 +datum=WGS84 +units=m +no_defs" # Robinson projection
world <- ne_countries(scale = "medium", returnclass = "sf") %>% 
  st_transform(cCRS)
Region <- c(xmin = 18, xmax = 130, ymin = -39, ymax = 34) # Indian Ocean limits
Bndry <- fSpatPlan_Get_Boundary(Region, cCRS) # boundary

@@

VentsInactive <- filter(Vents, Activity == "inactive")
VentsInactiveSF <- st_as_sf(VentsActive, coords = c(x = "Longitude", y = "Latitude"), crs = 4326) %>% 
  st_transform(cCRS)

@@

```{r vents plot, echo = FALSE, message = FALSE, warning = FALSE, results = "hide"}
world <- ne_countries(scale = "medium", returnclass = "sf")
IndianO <- c(xmin = 17.47941, ymin = -38.7402, xmax = 129.552, ymax = 33.82872)

ggplot() + 
  geom_sf(data = world) + 
  geom_sf(data = VentsActiveSF) +
  ggtitle("Active vents after st_tranform(cCRS)") +
  coord_sf(xlim = c(IndianO[["xmin"]], IndianO[["xmax"]]), ylim = c(IndianO[["ymin"]], IndianO[["ymax"]]), expand = FALSE)

#PUxVents <- st_join(PUs, VentsActiveSF, join = st_within) #no longer needed
#VentsActiveSFClean <- dplyr::select(VentsActiveSF, Name.ID, geometry)
#VentsActiveSFCleanBuff <- st_buffer(VentsActiveSFClean, dist = 1000)
#PUxVents <- st_intersects(VentsActiveSF, PUs)
#PUs <- PUs %>% 
#  mutate(Vents = PUxVents)

@@

ggplot() + 
  geom_sf(data = world) +
  geom_sf(data = NEIO_09,  aes(color="A"), show.legend = "line")+ 
  geom_sf(data = NWIO_14,  aes(color="A"), show.legend = "line")+
  geom_sf(data = SIO_11,  aes(color="A"), show.legend = "line")+
  geom_sf(data = SIO_19,  aes(color="A"), show.legend = "line")+
  geom_sf(data = SIO_22,  aes(color="A"), show.legend = "line")+
  geom_sf(data = SIO_23,  aes(color="A"), show.legend = "line")+
  geom_sf(data = SIO_30,  aes(color="A"), show.legend = "line")+
  geom_sf(data = SIO_32,  aes(color="A"), show.legend = "line")+
  geom_sf(data = SIO_35,  aes(color="A"), show.legend = "line")+
  geom_sf(data = SIO_36, aes(color="A"), show.legend = "line")+
  geom_sf(data = SIO_37,  aes(color="A"), show.legend = "line")+
  geom_sf(data = Seamounts,  aes(color="B"), show.legend = "line")+
  geom_sf(data = Plateaus,  aes(color="C"), show.legend = "line")+
  geom_sf(data = IBA,  aes(color="D"), show.legend = "line") + 
  geom_sf(data = IMMA,  aes(color="E"), show.legend = "line")+ 
  geom_sf(data = VentsActiveSF,  aes(color="F"), show.legend = "line")+
  geom_sf(data = VentsInactiveSF,  aes(color="G"), show.legend = "line")+
  ggtitle("IBA, IMMA, EBSAs, Seamounts, Plateaus, Vents")+ 
  coord_sf(xlim = c(IndianO[["xmin"]], IndianO[["xmax"]]), ylim = c(IndianO[["ymin"]], IndianO[["ymax"]]), expand = FALSE)+
  scale_color_manual(values = c("A" = "grey", "B" = "black", "C" = "blue", "D" = "green", "E" = "purple", "F" = "red", "G" = "brown"), 
                     labels = c("EBSAs", "Seamounts", "Plateaus", "IBAs", "IMMAs", "Active vents", "Inactive Vents"),
                     name = "Legend")

```
### Targets dataframe

```{r Conservation features and targets, echo = FALSE, message = FALSE, warning = FALSE}

Cons_feats <- tibble(Features = c("IBA", "IMMA", "NEIO_09", "NWIO_14", "SIO_11", "SIO_19", "SIO_22", "SIO_23", "SIO_30", "SIO_32", "SIO_35", "SIO_36", "SIO_37","Seamounts", "Plateaus", "Active_Vents", "Inactive_Vents"), Targets = c(0.3, 0.3, 0.3, 0.3, 0.3, 0.7, 0.7, 0.7, 0.7, 0.7, 0.7, 0.7, 0.7, 0.3, 0.3, 0.7, 0.7))

@@

+ PMN and CFC areas (Sharma, 2011 for USD valuation)
+ Exploration & reserved areas of deep-sea mining (lock out because won't agree to lose investments)
+ Fishing effort (using ready-made cost layer in Spatial Planning repo)
+ Shipping routes (use value of cell in Halpern et al. (2015), 2013 dataset for relative cost)

@@

### CFC

Cobalt-rich Ferromanganese Crusts. Still researching how to value.

```{r CFC, echo = FALSE, results = "hide", message = FALSE, warning = FALSE}
CFC <- "~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Costs/Mining_resources/Crust/Crust_areas_Hein2013.shp" 
PUxCFC <- fSpatPlan_Get_Polyg(CFC, PUs)  # function returns out as a logical vector
PUs <- PUs %>% 
  mutate(CFC = PUxCFC) # Add logical vector to variable
```

### PMN

Areas rich in PolyMetallic Nodules. **Monetary value** could be assigned using:

+ density: 5.6kg/sqm in Indian Ocean basin (Sharma, 2011) - an approximation but better than nothing?
[link](https://www.resolve.ngo/docs/mar_technol_soc_j_45_28a.pdf)
+ value: 320-1,100USD/tonne of nodules (CRU Consulting, 2019) [link](https://isa.org.jm/files/files/documents/CRU_ISA%20Polymetallic%20nodule%20valuation%20report_24Aug2020.pdf)
+ size of PMN area: #of PUs that intersect with PMN layer * PU size (1000sqkm)
+ extraction rate: 1.5Mt/yr for 20 yrs (Sharma, 2011) (but using this brings even greater uncertainty, I think)

```{r PMN, echo = TRUE,  results = "hide", message = FALSE, warning = FALSE}
PMN <- "~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Costs/Mining_resources/Nodules/Nodules_Hein2013.shp"
PUxPMN <- fSpatPlan_Get_Polyg(PMN, PUs)  
PUs <- PUs %>% 
  mutate(PMN = PUxPMN)
```

### Exploration areas

We need to **lock out** those areas, as mining companies have invested heavily into exploration and will not agree to lose access.

```{r expl, echo = FALSE, results = "hide", message = FALSE, warning = FALSE}
ExplPMN <- "~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Costs/Mining_contractors/ContractorAreas/01_pmn_exploration_areas.shp"
PUxExplPMN <- fSpatPlan_Get_Polyg(ExplPMN, PUs)  
PUs <- PUs %>% 
  mutate(ExplPMN = PUxExplPMN)

ExplCFC <- "~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Costs/Mining_contractors/ContractorAreas/03_cfc_exploration_areas.shp"
PUxExplCFC <- fSpatPlan_Get_Polyg(ExplCFC, PUs)  
PUs <- PUs %>% 
  mutate(ExplCFC = PUxExplCFC)
```

### Reserved areas

Areas reserved for mining exploitation by developing countries - can also be locked out (same reasoning as above).

```{r res, echo = FALSE, results = "hide", message = FALSE, warning = FALSE}
ResPMN <- "~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Costs/Mining_contractors/ContractorAreas/04_pmn_reserved_areas.shp"
PUxResPMN <- fSpatPlan_Get_Polyg(ResPMN, PUs)  
PUs <- PUs %>% 
  mutate(ResPMN = PUxResPMN)

ResCFC <- "~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Costs/Mining_contractors/ContractorAreas/05_cfc_reserved_areas.shp"
PUxResCFC <- fSpatPlan_Get_Polyg(ResCFC, PUs)  
PUs <- PUs %>% 
  mutate(ResCFC = PUxResCFC)
```

```{r Mining layer, echo = FALSE, message = FALSE, warning = FALSE, results = "hide"}
world <- ne_countries(scale = "medium", returnclass = "sf")
IndianO <- c(xmin = 17.47941, ymin = -38.7402, xmax = 129.552, ymax = 33.82872)
CFC <- st_read("~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Costs/Mining_resources/Crust/Crust_areas_Hein2013.shp")
PMN <- st_read ("~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Costs/Mining_resources/Nodules/Nodules_Hein2013.shp")
Expl_PMN <- st_read ("~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Costs/Mining_contractors/ContractorAreas/01_pmn_exploration_areas.shp")
Expl_PMS <- st_read ("~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Costs/Mining_contractors/ContractorAreas/02_pms_exploration_areas.shp")
Expl_CFC <- st_read ("~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Costs/Mining_contractors/ContractorAreas/03_cfc_exploration_areas.shp")
Res_PMN <- st_read ("~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Costs/Mining_contractors/ContractorAreas/04_pmn_reserved_areas.shp")
Res_CFC <- st_read ("~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Costs/Mining_contractors/ContractorAreas/05_cfc_reserved_areas.shp")
```

Mining areas

>>>>

---
title: "Cross-sectoral conservation of the Indian Ocean"
author: "Lea Fourchault"
date: "08/03/2022"
output: 
 html_document:
    toc: true
    toc_float: true

---
<style type="text/css">

@@

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
---

+ how to group all files (ex: all EBSAs + IBAs + IMMAs + vents + seamounts...) into one, to create ConsFeatures? Or is it enough to have columns next to PUs? RESOLVED: yes, stacking = creating columns next to PUs.
+ should we use Sharma (2011) + CRU Consulting (2019) to assign USD values to PMN areas? see PMN subsection. RESOLVED, yes, can use for this project (need to cross-check if want to publish)
+ shipping raster : cells have certain values (# of ships recorded?? I can't figure from the attached files, metadata is confusing), could allow to infer USD value of route (# of ships * approx cost of fuel for average cargo to go through 1 PU, which is 1000sqkm)? So we wouldn't try to compute the cost of deviating from any given route, but rather compute the price for each ship to go through 1 PU * the number of ships going through that PU. It sounds fairly logical to me, but maybe it's not? RESOLVED: check 2008 metada, run prio problem with only shipping
+ fishing cost layer : there is a ready-made fx in SpatPlan repo (thanks Jase & Tin), should we use it for this project? see fishing subsection (but issues with plotting  as you can see). RESOLVED: yes, use fx from Jase & Tin. Change scale to remove outliers/crop to planning region.
+ I'm using echo = FALSE, message = FALSE, warning = FALSE, results = "hide" and still can't get rid of some coding output in the .html. RESOLVED: cut up chuncks of code with separate ones for plots
+ I st_to_sf the vents from .csv with PUs' crs but have some issues intersecting the points with polygs, will look more into it. RESOLVED: use st_contains after crs = 4326 & st_transform(cCRS)
+ add list of species as cons_feats: EN, VU etc (Ant's email to mining colleague)? RESOLVED: fine-tune later
+ shipping raster (Halpern et al., 2015) is too heavy to be converted to polygon, what could be other methods (not using as.polygon)? RESAMPLE if resolution <1°. Resolution = 934m*934m (! 1° at equator = 111km = 60 nm; [convert](https://www.johndcook.com/how_big_is_a_degree.html); resolution too fine so need resampling). RESOLVED: no resampling, overlay raster with PUs (exactextractr) to attribute mean raster cells value to PUs (no weighted mean because raster cells are all equal-area because Mollweide proj)
+ how to remove EEZs from PUs? RESOLVED: find EEZ shp and spatial predicates, inverse overlay with Bndry, recreate PUs
+ setting  targets. RESOLVED: 30% for rpz area, 70% for important/rare feature; create dataframe with 1 column: features, 1 column: targets (match order)
+ *cross inactive vents & PUs - done*
+ check research questions. RESOLVED keep Ant's questions
+ Error in h(simpleError(msg, call)) : error in evaluating the argument 'x' in selecting a method for function 'add_relative_targets': error in evaluating the argument 'features' in selecting a method for function 'problem': object 'Cons_feats' not found. RESOLVED: use string: features = "x", "y"; instead of creating Cons_feats tibble
+ cannot install lpsymphony: almost compiled but failed because did not find directory for gfortran AND see this link: issues with Mojave https://www.cynkra.com/blog/2021-03-16-gfortran-macos/ : brew and gfortran cask deprecated since 2021 (still work if installed before but i didn't). RESOLVED: install Gurobi using student email
+ PUs: when mutating FishingCost column to PUs, creates duplicate of 'geometry' column called 'Fishing.geometry'. RESOLVED don't forget to add dataframe: dataframeDOLLARcolumn name: PUs <- PUs %>% mutate(Fishing = FishingCost$FishingCost)
+ *create new PUs with ABNJ only - done but check method*, methodology not checked
+ general bugs to run code on .rmd file. Use knitr::spin instead. I also replaced the initial chunk with ```{r setup, include=FALSE} knitr::opts_chunk$set(echo = TRUE)``` but no result, maybe need to reinstall r and r studio

@@

Current questions & updates:

+ ongoing - pricing deep-sea mineral resources: for CFC *we know the area based on Crust.shp; we know CFC mean specific-surface areas = 325 m2/g (Hein et al., 2000; found in Hein et al., 2013) -> can find tonnage; then, we know chemical composition of CFC in Indian Ocean (given in wt% or ppm=g/tonne, for ex: 392 Mo ppm; Fe=22wt% - see Hein, 2014 or Petersen et al., 2016) -> can multiply by CFC tonnage to get tonnage for each mineral; then multiply by market prices (mean over 5 yrs?) to estimate total value of CFC in Indian Ocean?*
+ I have emailed Dr. Mizell about her calculations of CFC tonnage for the Indian Ocean (paper only gives global tonnage)

*Coding questions, will think about it some more*

+ how to assign price to mining shps / create mining cost layer once value of deep-sea resources is computed?
+ overlay shipping & PUs + plot with log10: partly resolved using exact_extractr but issues with +1 and log10 and when plotting PUxShipping: not a map but a scatterplot; issues with transformation/reprojection?

@@

+ PUs = ABNJ only
+ cons_feats = IMMAs, IBAs, chosen EBSAs, vents (active + inactive), seamounts, plateaus; no aquamaps
+ PMN mining value = value/t * density * area, discount rate and other improvements to be determined later (if any), USD
+ fishing cost = removed fish mass * fish value (layer created by Jase), USD
+ shipping cost = mean of shipping raster cells' values for each PU using exact_extracr, not USD (intensity) OR SHOULD WE USE SHIPPING AS A PENALTY?
+ targets = 30% for rpz area, 70% for important/rare feature

@@

**Aim**

@@

**Current research questions**

1. How can future seabed mining be included in spatial planning?
2. What is the potential spatial overlap among biodiversity, fishing, shipping and seabed mining in the Indian Ocean high seas?
3. Where could we place reserves in the Indian Ocean high seas that meet conservation objectives whilst minimsing conflict with fishing and seabed mining?

@@

blue habitats do not include hydro vents, so will use Beaulieu & Szanfranski (2020) [link](https://doi.pangaea.de/10.1594/PANGAEA.917894?format=html#download)

@@

cCRS <- "+proj=robin +lon_0=0 +x_0=0 +y_0=0 +datum=WGS84 +units=m +no_defs" # Robinson projection
world <- ne_countries(scale = "medium", returnclass = "sf") %>% 
  st_transform(cCRS)
#Region <- c(xmin = 18, xmax = 140, ymin = -39, ymax = 34) 
Region <- c(xmin = 20, xmax = 120, ymin = -60, ymax = 20) 
# Indian Ocean limits: CIA World Factbook (but see: According to the International Hydrographic Organization (IHO) and the United Nations Oceans Atlas, the area from 40°S latitude to 60°S latitude is included in the Indian Ocean. The area that encircles the globe from 60°S latitude to the coast of Antarctica is called The Great Southern Ocean. The Indian Ocean's width extends from 45° E longitude to 110°East longitude.) Also, Region has big impact on final plan because fishing cost is much lower as latitude increases towards south pole
Bndry <- fSpatPlan_Get_Boundary(Region, cCRS) # boundary

@@

VentsInactive <- filter(Vents, Activity == "inactive")
VentsInactiveSF <- st_as_sf(VentsInactive, coords = c(x = "Longitude", y = "Latitude"), crs = 4326) %>% 
  st_transform(cCRS)

@@

```{r vents plot, echo = FALSE, message = FALSE, warning = FALSE, results = "hide"}
world <- ne_countries(scale = "medium", returnclass = "sf")
#IndianO <- c(xmin = 17.47941, ymin = -38.7402, xmax = 129.552, ymax = 33.82872)
IndianO <- c(xmin = 20, xmax = 120, ymin = -60, ymax = 20)

#ggplot() + 
 # geom_sf(data = world) + 
  #geom_sf(data = VentsInactiveSF) +
  #ggtitle("Active vents after st_tranform(cCRS)") +
  #coord_sf(xlim = c(IndianO[["xmin"]], IndianO[["xmax"]]), ylim = c(IndianO[["ymin"]], IndianO[["ymax"]]), expand = FALSE)

# no longer needed
#PUxVents <- st_join(PUs, VentsActiveSF, join = st_within) 
#VentsActiveSFClean <- dplyr::select(VentsActiveSF, Name.ID, geometry)
#VentsActiveSFCleanBuff <- st_buffer(VentsActiveSFClean, dist = 1000)
#PUxVents <- st_intersects(VentsActiveSF, PUs)
#PUs <- PUs %>% 
#  mutate(Vents = PUxVents)

@@

IndianO <- c(xmin = 20, xmax = 120, ymin = -60, ymax = 20)

ggplot() + 
  geom_sf(data = world) +
  geom_sf(data = NEIO_09,  aes(color="A"), show.legend = "line")+ 
  geom_sf(data = NWIO_14,  aes(color="A"), show.legend = "line")+
  geom_sf(data = SIO_11,  aes(color="A"), show.legend = "line")+
  geom_sf(data = SIO_19,  aes(color="A"), show.legend = "line")+
  geom_sf(data = SIO_22,  aes(color="A"), show.legend = "line")+
  geom_sf(data = SIO_23,  aes(color="A"), show.legend = "line")+
  geom_sf(data = SIO_30,  aes(color="A"), show.legend = "line")+
  geom_sf(data = SIO_32,  aes(color="A"), show.legend = "line")+
  geom_sf(data = SIO_35,  aes(color="A"), show.legend = "line")+
  geom_sf(data = SIO_36, aes(color="A"), show.legend = "line")+
  geom_sf(data = SIO_37,  aes(color="A"), show.legend = "line")+
  geom_sf(data = Seamounts,  aes(color="B"), show.legend = "line")+
  geom_sf(data = Plateaus,  aes(color="C"), show.legend = "line")+
  geom_sf(data = IBA,  aes(color="D"), show.legend = "line") + 
  geom_sf(data = IMMA,  aes(color="E"), show.legend = "line")+ 
  geom_sf(data = VentsActiveSF,  aes(color="F"), show.legend = "line")+
  geom_sf(data = VentsInactiveSF,  aes(color="G"), show.legend = "line")+
  ggtitle("IBA, IMMA, EBSAs, Seamounts, Plateaus, Vents")+ 
  coord_sf(xlim = c(18, 130), ylim = c(22, -60), expand = FALSE)+
  scale_color_manual(values = c("A" = "grey", "B" = "black", "C" = "blue", "D" = "green", "E" = "purple", "F" = "red", "G" = "brown"), 
                     labels = c("EBSAs", "Seamounts", "Plateaus", "IBAs", "IMMAs", "Active vents", "Inactive Vents"),
                     name = "Legend")

### Longhurst provinces

```{r longhurst plot, echo = FALSE, message = FALSE, warning = FALSE}

longh <- st_read(file.path("Data","LonghurstProvinces","Longhurst_world_v4_2010.shp")) %>% 
  st_make_valid()
longhRob <- st_transform(longh, cCRS)

nr <- st_nearest_feature(PUs, longhRob)

PUs <- PUs %>% 
  mutate(ProvCode = longh$ProvCode[nr],
         ProvDescr = longh$ProvDescr[nr])

ggplot() + 
  geom_sf(data = world, color = "grey20", fill = "grey20", size = 0.1, show.legend = FALSE) +
  geom_sf(data = PUs, aes(fill = ProvCode), colour = "grey50", size = 0.05, alpha = 0.7, show.legend = TRUE) +
  coord_sf(xlim = c(18, 130), ylim = c(22, -60), expand = FALSE)

### Targets dataframe

#Cons_feats <- tibble(Features = c("IBA", "IMMA", "NEIO_09", "NWIO_14", "SIO_11", "SIO_19", "SIO_22", "SIO_23", "SIO_30", "SIO_32", "SIO_35", "SIO_36", "SIO_37","Seamounts", "Plateaus", "Active_Vents", "Inactive_Vents"), Targets = c(0.3, 0.3, 0.3, 0.3, 0.3, 0.7, 0.7, 0.7, 0.7, 0.7, 0.7, 0.7, 0.7, 0.3, 0.3, 0.7, 0.7))

## Locked out areas

### polymetallic nodules / polymetallic sulfides: exploration & reserved areas

```{r Locked out, echo = FALSE, message = FALSE, warning = FALSE, results = "hide"}

ExplPMN <- "~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Costs/Mining_contractors/ContractorAreas/01_pmn_exploration_areas.shp"
PUxExplPMN <- fSpatPlan_Get_Polyg(ExplPMN, PUs)  
PUs <- PUs %>% 
  mutate(ExplPMN = PUxExplPMN)

ResPMN <- "~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Costs/Mining_contractors/ContractorAreas/04_pmn_reserved_areas.shp"
PUxResPMN <- fSpatPlan_Get_Polyg(ResPMN, PUs)  
PUs <- PUs %>% 
  mutate(ResPMN = PUxResPMN)

```{r Plotting Locked out, echo = FALSE, message = FALSE, warning = FALSE}
# plotting locked out areas

world <- ne_countries(scale = "medium", returnclass = "sf")
IndianO <- c(xmin = 18, ymin = -60, xmax = 130, ymax = 22)
Expl_PMN <- st_read ("~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Costs/Mining_contractors/ContractorAreas/01_pmn_exploration_areas.shp")
Expl_PMS <- st_read ("~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Costs/Mining_contractors/ContractorAreas/02_pms_exploration_areas.shp")
Res_PMN <- st_read ("~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Costs/Mining_contractors/ContractorAreas/04_pmn_reserved_areas.shp")

ggplot() + 
  geom_sf(data = world) +
  geom_sf(data = Expl_PMN, aes(color="A"), show.legend = "line") + 
  geom_sf(data = Expl_PMS, aes(color="B"), show.legend = "line") +
  geom_sf(data = Res_PMN, aes(color="C"), show.legend = "line") +
  scale_color_manual(values = c("A" = "red", "B" = "yellow", "C" = "black"),
                     labels = c("Expl_PMN", "Expl_PMS", "Res_PMN"),
                     name = "Legend") +
  ggtitle("Exploration & Reserved areas") + 
  coord_sf(xlim = c(18, 130), ylim = c(22, -60), expand = FALSE)

@@

+ PMN and CFC areas (Sharma, 2011/ Mizell et al., 2022 for USD valuation) - in progress
+ Fishing effort (using ready-made cost layer in Spatial Planning repo)
+ Shipping routes (use value of cell in Halpern et al. (2015), 2013 dataset for relative cost) - in progress, but should be used as penalty, not cost?

@@

### Mineral resources

```{r Plot PMN CFC, echo = FALSE, message = FALSE, warning = FALSE}
#' Plotting mining resources

world <- ne_countries(scale = "medium", returnclass = "sf")
#IndianO <- c(xmin = 17.47941, ymin = -38.7402, xmax = 129.552, ymax = 33.82872)
IndianO <- c(xmin = 20, xmax = 120, ymin = -60, ymax = 20)
CFC <- st_read("~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Costs/Mining_resources/Crust/Crust_areas_Hein2013.shp")
PMN <- st_read ("~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Costs/Mining_resources/Nodules/Nodules_Hein2013.shp")
ggplot() + 
  geom_sf(data = world) +
  geom_sf(data = CFC, aes(color="A"), show.legend = "line") + 
  geom_sf(data = PMN, aes(color="B"), show.legend = "line") + 
  scale_color_manual(values = c("A" = "grey", "B" = "brown"),
                     labels = c("CFC", "PMN"),
                     name = "Legend") +
  ggtitle("Polymetallic Nodules and Cobalt-rich Ferromanganese Crust areas") + 
  coord_sf(xlim = c(18, 130), ylim = c(22, -60), expand = FALSE)

CFC value
script to get area modified from https://rpubs.com/rural_gis/255550

```{r CFC, echo = FALSE, message = FALSE, warning = FALSE, results = "hide"}
CFC <- "~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Costs/Mining_resources/Crust/Crust_areas_Hein2013.shp" 
PUxCFC <- fSpatPlan_Get_Polyg(CFC, PUs)  # function returns out as a logical vector
PUs <- PUs %>% 
  mutate(CFC = PUxCFC)

#load CFC areas polygon shapefile 
CFC <- st_read("~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Costs/Mining_resources/Crust/Crust_areas_Hein2013.shp")
CFCRobin <- st_transform(CFC, cCRS) #robin proj

#run the intersect function, converting the output to a tibble in the process
#int <- as_tibble(st_intersection(CFCRobin, BndryABNJ))
int <- st_intersection(CFCRobin, BndryABNJ)

#add in an area count column to the tibble (area of each CFC poly in intersect layer)
int$CFCarea <- st_area(int)
```

Intersections between PUs and CFC in red - correct

```{r plot CFC, echo = FALSE, message = FALSE, warning = FALSE}

#plot the layers to visually check result of intersect
plot (CFCRobin$geometry, col='green')
plot(BndryABNJ$geometry, add=T)
plot(int, col='red', add=T) #correct

#group data by county area and calculate the total arable land area per county
#output as new tibble
AreaCFCtotal <- int %>%
  summarise(areaCFCtot = sum(CFCarea))
```

```
# CFC value code (so you can have a look)
# script to get area modified from https://rpubs.com/rural_gis/255550

CFC <- "~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Costs/Mining_resources/Crust/Crust_areas_Hein2013.shp" 
PUxCFC <- fSpatPlan_Get_Polyg(CFC, PUs)  # function returns out as a logical vector
PUs <- PUs %>% 
  mutate(CFC = PUxCFC)

#load CFC areas polygon shapefile 
CFC <- st_read("~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Costs/Mining_resources/Crust/Crust_areas_Hein2013.shp")
CFCRobin <- st_transform(CFC, cCRS) #robin proj

#run the intersect function, converting the output to a tibble in the process
#int <- as_tibble(st_intersection(CFCRobin, BndryABNJ))
int <- st_intersection(CFCRobin, BndryABNJ)

#add in an area count column to the tibble (area of each CFC poly in intersect layer)
int$CFCarea <- st_area(int)

#plot the layers to visually check result of intersect
plot (CFCRobin$geometry, col='green')
plot(BndryABNJ$geometry, add=T)
plot(int, col='red', add=T) #correct

#group data by county area and calculate the total arable land area per county
#output as new tibble
AreaCFCtotal <- int %>%
  summarise(areaCFCtot = sum(CFCarea)) # = 2.256596e+12 [m^2]
```
**CFC area = 2.256596e+12 [m^2], seems a lot (see code above, no error messages but probably wrong somewhere)? Would be 2256596 sqkm, and I have 39803 PUs of 1000 sqkm each**

*Other method to check :* 
area of CFC in PUs < 2753*1000 sqkm = 2.753e+6 sqm (see sum of TRUE intersections between PUs and CFC poly below)

```
sum(PUxCFC) # sum of cross = TRUE between PUs and CFC polyg
[1] 2753

```
PMN - same code

```{r PMN, echo = FALSE, message = FALSE, warning = FALSE, results = "hide"}

# PMN value

#script to get area modified from https://rpubs.com/rural_gis/255550

#load county areas polygon shapefile

PMN <- "~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Costs/Mining_resources/Nodules/Nodules_Hein2013.shp"
PUxPMN <- fSpatPlan_Get_Polyg(PMN, PUs)  
PUs <- PUs %>% 
  mutate(PMN = PUxPMN)

PMN <- st_read("~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Costs/Mining_resources/Nodules/Nodules_Hein2013.shp")
PMNRobin <- st_transform(PMN, cCRS) #robin proj

#run the intersect function, converting the output to a tibble in the process
int2 <- st_intersection(PMNRobin, BndryABNJ) #need this to avoid Error in UseMethod("st_area") : 
#int2 <- as_tibble(st_intersection(PMNRobin, BndryABNJ)) #no applicable method for 'st_area' applied to an object of class "c('tbl_df', 'tbl', 'data.frame')"

#add in an area count column to the tibble (area of each CFC poly in intersect layer)
int2$PMNarea <- st_area(int2)
```

Intersection between PUs and PMN in red - correct

```{r plot PMN, echo = FALSE, message = FALSE, warning = FALSE}

#plot the layers to visually check result of intersect
plot (PMNRobin$geometry, col='green')
plot(BndryABNJ$geometry, add=T)
plot(int2, col='red', add=T) #correct
